A Decision Procedure for (Co)datatypes in SMT Solvers

### 2015-02-20 tex/body.tex
more

%\ourparagraph{Integration into {\relax\XXXL($T\negvthinspace$)}.}
{\looseness=-1
The calculus is implemented as a \relax{theory solver} of CVC4, that is, a
specialized solver for determining the satisfiability of conjunctions of
constraints for its theory.
Given a theory $T = T_1 \mathrel\cup \cdots \mathrel\cup T_n$ and a set of
input clauses $F$ in conjunctive normal form, the \XXXL($T$) procedure
incrementally builds partial assignments from the atomic formulas of $F$ to
truth values such that no clause in $F$ is falsified. We can regard such a
partial assignment as a set $M$ of true literals.
% where $a \in M$ for all atoms $a$ that
%it assigned the value $\top$ and $\neg\; a \in M$ for all atoms that it assigns
%the value $\bot$.
%at a high level,
By a variant
\cite{jovanovic2011sharing} of the Nelson--Oppen method,
each $T_i$-solver %for $i = 1, \ldots, l$  n?
takes as input the union $M_i$
of (1)~the purified form of $T_i$-constraints occurring in $M$, where terms
containing symbols not belonging to $T_i$ are replaced by fresh variables\begin{rep} of the
same type\end{rep}; (2)~additional (dis)equalities between variables of
types not belonging to $T_i$.
Each $T_i$-solver
either reports that a subset $C$ of $M_i$ is $T_i$-unsatisfiable, in which case
$\neg\; C$ is added to $F\!$, adds a clause to $F\!$, or does nothing.
When $M$ is a complete assignment for $F\!$, a theory solver can choose to do
nothing only if $M_i$ is indeed $T_i$-satisfiable.

>>>>

%\ourparagraph{Integration into {\relax\XXXL($T\negvthinspace$)}.}
{\looseness=-1
The decision procedure is implemented as a \relax{theory solver} of CVC4, that is, a
specialized solver for determining the satisfiability of conjunctions of
constraints for its theory.
Given a theory $T = T_1 \mathrel\cup \cdots \mathrel\cup T_n$ and a set of
input clauses $F$ in conjunctive normal form, the \XXXL($T$) procedure
incrementally builds partial assignments of
truth values to the atoms of $F$ such that no clause in $F$ is falsified. We can regard such a
partial assignment as a set $M$ of true literals.
% where $a \in M$ for all atoms $a$ that
%it assigned the value $\top$ and $\neg\; a \in M$ for all atoms that it assigns
%the value $\bot$.
%at a high level,
By a variant of the Nelson--Oppen method \cite{jovanovic2011sharing,nelson-oppen-1979},
each $T_i$-solver %for $i = 1, \ldots, l$  n?
takes as input the union $M_i$
of (1)~the purified form of $T_i$-constraints occurring in $M$, where fresh
variables replace terms containing symbols not belonging to~$T_i$;
(2)~additional (dis)equalities between variables of types not belonging to
$T_i$. Each $T_i$-solver
either reports that a subset $C$ of $M_i$ is $T_i$-unsatisfiable, in which case
$\neg\; C$ is added to $F\!$, adds a clause to $F\!$, or does nothing.
When $M$ is a complete assignment for $F\!$, a theory solver can choose to do
nothing only if $M_i$ is indeed $T_i$-satisfiable.
